---
title: "Assignment 3 Task 4"
author: "Will Geiken"
date: "10/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(kableExtra)
library(tidyverse)
library(janitor)
library(here)

urchin_data <- read.csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Introduction

In this document I will be 

### 1. Wrangling

Wrangle the data such that it only includes Naples Reef observations of purple urchins from January 2010.

```{r, echo = TRUE, message = FALSE, warning = FALSE}
purp_urch_jan_2010 <- urchin_data %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>%
  mutate(month = month(date)) %>% 
  filter(month == 1, year == 2010, site == "NAPL")
```

### 2. Visualization

Create a histogram and Q-Q plot comparing the two treatment types 

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(data = purp_urch_jan_2010, aes(x = size)) +
  geom_histogram(bins = 10) +
  facet_wrap(~treatment)
```
Takeaways:

Create a Q-Q plot comparing the two treament types

```{r, echo = TRUE, message = FALSE, warning = FALSE}
ggplot(data = purp_urch_jan_2010, aes(sample = size)) +
  geom_qq() +
  facet_wrap(~treatment)
```
Takeaways:

### Data table

Create a data table that displays descriptive statistics for each group. Include mean, standard deviation, and sample size

```{r, echo = TRUE, message = FALSE, warning = FALSE}
urchin_table <- purp_urch_jan_2010 %>% 
  group_by(treatment) %>% 
  summarise("Mean Diameter (cm)" = mean(size),
            "Median Diameter (cm)" = median(size),
            "Standard Deviation (cm)" = sd(size),
            "Sample Size" = n()) 

kbl(urchin_table) %>% 
  kable_styling()
```

### Perform a two-sample t-test to compare means between the two groups

```{r, echo = TRUE, message = FALSE, warning = FALSE}
annual_sample <- purp_urch_jan_2010 %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_sample <- purp_urch_jan_2010 %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size)

t.test(x = annual_sample, y = control_sample)
```

### Discussion of Results


### Meaning of the p-value
